% =============================================================================
% DATASET + CONFOUNDERS   --   target ~1.25 pp. Borrowed from paper/ and cut.
% Cuts vs paper/ (logged in docs/faimi/dataset-section-cuts.md): dropped
% tab:notation (symbols inlined), fig:volumes, tab:volumes (folded to one
% sentence), the connected-components paragraph, the GE/Siemens voxel arithmetic
% (-> 1 line). KEPT: tab:cohort (fig:anatomy dropped for 8pp). Added: gold/silver comparability
% (load-bearing for E2) -- still needs an empirical one-liner.
% =============================================================================
\section{Dataset and Exploratory Analysis}
\label{sec:dataset}

\subsection{The CSpineSeg Cohort}
\label{sec:cohort}

We work with CSpineSeg~\cite{zhou2025cspineseg}, a public collection of 1,255
sagittal T2-weighted cervical-spine MRI exams from 1,232 patients at Duke
University. We drop a single localizer series with anomalous dimensions, leaving a
working set $\mathcal{D}_0$ of $N_0 = 1{,}254$ exams from 1,231 patients. The task
is semantic segmentation: labelling every voxel of a scan as background, vertebral
body, or intervertebral disc ($\mathcal{L} = \{0, 1, 2\}$).
About $40\%$ of the exams carry expert (gold) reference segmentations, drafted by one post-doctoral researcher without medical training and reviewed and revised by
one of six board-certified radiologists; the rest were segmented automatically by the authors' own model, and
these silver masks were never human-reviewed -- though the mid-sagittal slice of
each unannotated scan was checked for image quality.

We write the cohort as $\mathcal{D} = \{(X_i, Y_i, A_i, q_i)\}_{i=1}^{N}$: exam
$i$ pairs a 3D image $X_i$ with a reference segmentation $Y_i$, an attribute value
$A_i$ over the protected attributes $\mathcal{A} = \{\text{sex}, \text{race},
\text{age}\}$, and a provenance flag $q_i \in \{\mathsf{g}, \mathsf{s}\}$ marking a
gold (expert) or silver (machine-generated) label. We write
$\mathcal{D}_{A=a}$ for the subgroup with $A = a$. The analysis cohort
$\mathcal{D}$ ($N = 1{,}142$) is the sex-balanced subset of $\mathcal{D}_0$ defined
in ``\nameref{sec:regimes}''.

\autoref{tab:cohort} summarises $\mathcal{D}_0$: it leans slightly female, is
predominantly White with a sizeable Black minority, overwhelmingly non-Hispanic,
spans the adult age range, and splits across two manufacturers and two field
strengths. The volumes are strongly anisotropic -- high in-plane resolution
(${\sim}0.53$\,mm) but thick, few sagittal slices (${\sim}4$\,mm, 12--25 per scan).
These group sizes decide which axes we test formally -- sex (F/M), race
(White/Black), and age have enough exams per group for reliable inference --
while the smaller race categories (8\% combined) and ethnicity are too sparse and
reported descriptively only.

\begin{table}[t]
\centering
\caption{Composition of the working set $\mathcal{D}_0$ ($N_0 = 1{,}254$ exams,
counted at the exam level). Percentages are of $N_0$; age is mean\,$\pm$\,SD and the
13 missing-age exams (confirmed ${>}89$) are grouped into $\ge$60. The cohort is
overwhelmingly non-Hispanic ($92.7\%$); the $7.7\%$ Other/unknown race is mostly Asian
and unreported.}
\label{tab:cohort}
\small
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r l r}
\hline\noalign{\smallskip}
Characteristic & $n$ (\%) & Characteristic & $n$ (\%) \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\textit{Sex} & & \textit{Age, years} & $54.6 \pm 16.3$ \\
\quad Female & 683 (54.5) & \quad $<$40 & 241 (19.2) \\
\quad Male & 571 (45.5) & \quad 40--60 & 486 (38.8) \\
\textit{Race} & & \quad $\ge$60 & 527 (42.0) \\
\quad White & 809 (64.5) & \textit{Manufacturer} & \\
\quad Black & 349 (27.8) & \quad Siemens & 788 (62.8) \\
\quad Other/unknown & 96 (7.7) & \quad GE & 466 (37.2) \\
\textit{Field strength} & & & \\
\quad 1.5\,T & 746 (59.5) & & \\
\quad 3.0\,T & 508 (40.5) & & \\
\noalign{\smallskip}\hline
\end{tabular*}
\end{table}

\subsection{Anatomical Confounders}
\label{sec:confounders}

A performance gap is bias only if anatomy does not explain it, and structure size is the
main confound: men's vertebral bodies and discs are ${\sim}23$--$24\%$ larger, and the
vertebral body also grows with age as cervical degeneration
accrues~\cite{kato2012cervical} and differs by race, while the disc stays essentially
constant -- the least size-confounded cross-check. We therefore report vertebral body and
disc separately alongside their macro average, and score with a volume-aware nDSC
(``\nameref{sec:metrics}'') beside raw Dice so a size effect is not read as bias. The
size gap is anatomy, not scanner (GE vs.\ Siemens differ ${\sim}2.15\times$ in voxel count
yet ${\sim}6\%$ in mm$^3$), and no attribute is entangled with the scanner; only age is
mildly entangled with race (White patients slightly older), so we control for age when
comparing race.
