% =============================================================================
% INTRODUCTION + (folded) RELATED WORK   --   target ~1.25 pp incl. contributions
% No standalone Related Work section at 8 pp: positioning folded into para 4.
% Framing: lineage-first (Parikh et al.), leakage-forward findings (Aasa's story):
% contribution 2 is the ~8-pt leakage/overestimation; contribution 3 is the
% fairness-specific false-confidence mode that the same leakage produces.
% =============================================================================
\section{Introduction}

Automated segmentation of the cervical spine -- labelling the vertebral bodies and
intervertebral discs on MRI -- helps clinicians grade stenosis and track degeneration
over time~\cite{kato2012cervical}. A model that works less well for some patient groups
than for others would quietly widen existing gaps in spine care, and bias audits are
increasingly required before clinical use, not an optional extra. Yet no demographic
fairness audit exists for this anatomy.

Auditing a segmentation model is harder than auditing a classifier. Its output is
millions of voxels rather than a single decision, so fairness has to be reduced to one
score (Dice, HD95) -- and that score is itself biased by structure size when anatomy
differs between groups, because the same boundary error counts for less on a larger
structure. Expert annotation is also costly, so datasets increasingly mix a few expert
(gold) labels with many machine-generated (silver) ones. If the silver labels are
flawed, the reference used to judge a model is itself unreliable.

Parikh et al.~\cite{parikh2025labelbias} introduced this ``biased ruler'' effect for
segmentation in breast MRI: a biased silver reference can \emph{inflate} apparent
disparity (``false magnitude''). We do not claim this effect; we take their
framework~\cite{parikh2025labelbias,parikh2025whofails} to a new anatomy and find a
second way it can mislead.

Fairness in segmentation is much less studied than in classification. The only prior
spine fairness study, FairMedFM~\cite{jin2024fairmedfm}, covers the lumbar spine, sex
only, and uses no volume correction; FairSeg~\cite{tian2024fairseg} is in ophthalmology.
None audit cervical-spine MRI across several protected attributes, and none ask whether
where the labels come from -- for training or for evaluation -- can by itself flip a
fairness verdict here.

\smallskip\noindent\textbf{Contributions.}\enspace Building on the biased-ruler
framework of Parikh et al.~\cite{parikh2025labelbias,parikh2025whofails}, we contribute:
(1)~the \textbf{first demographic fairness audit of cervical-spine MRI segmentation}
(CSpineSeg, $1{,}254$ exams) across sex, race, and age, on which the deployed model is
fair by every standard metric; (2)~evidence that a dataset's machine-generated labels
\textbf{leak information from the expert labels they were derived from}, so scoring
identical predictions against silver rather than expert labels \textbf{overestimates
performance by ${\sim}8$ Dice points} -- which affects \emph{any} user who validates on
such a dataset, not only fairness auditors; (3)~a \textbf{complementary mode of the
biased-ruler effect} from that same leakage -- the machine ruler flips age from
non-significant to significant not by widening the gap (Parikh et al.'s \emph{false
magnitude}) but by collapsing within-group variance (\emph{false confidence}); which mode
shows up depends on whether the silver ruler is independently biased (theirs) or a
near-copy of the model (ours); and (4)~\textbf{practitioner guidance} -- report
performance and fairness against expert labels, and treat a reference that scores both
much higher and much more tightly than experts (${\sim}4\times$ variance collapse) as too
close to the model to be an independent benchmark.
