% =============================================================================
% METHODOLOGY   --   target ~1.5 pp. Compressed from paper/sections/methodology.tex.
% 2 subsections; finer beats are \paragraph{} run-ins. One display eq (nDSC);
% Dice + DPD/DIR inline. Floats: fig:regimes (partition), tab:experiments. Cuts in
% docs/faimi/content-cut-from-paper.md.
% =============================================================================
\section{Methodology}
\label{sec:methodology}

Our design carries the label-bias and biased-ruler framework of Parikh et
al.~\cite{parikh2025labelbias,parikh2025whofails} from breast DCE-MRI over to
cervical-spine segmentation, with one structural difference that shapes everything
below: each CSpineSeg exam carries a gold or a silver label, never both.

\subsection{Data, Splits, and Label Regimes}
\label{sec:regimes}

The provenance flag $q_i$ partitions the cohort into disjoint gold and silver
sets, $\mathcal{D}^{\mathsf{g}}$ and $\mathcal{D}^{\mathsf{s}}$
($\mathcal{D}^{\mathsf{g}} \cap \mathcal{D}^{\mathsf{s}} = \varnothing$). We split
the cohort at the patient level into training, validation, and test sets in a
$70/10/20$ ratio, stratified by the protected attributes $\mathcal{A}$; splitting
by patient keeps
all of a multi-exam patient's exams on one side, preventing leakage. We then
balance the sexes -- the working set leans female (``\nameref{sec:cohort}''), and
an unequal base rate would conflate a true performance gap with mere
representation -- by randomly downsampling female exams within each split until the
two are equal, dropping $112$ exams ($N_0 = 1{,}254 \to N = 1{,}142$); this removes
representation as an explanation rather than attempting to close the
gap~\cite{parikh2025labelbias}. Because nnU-Net is cross-validation-native, we pool
$\mathcal{D}_{\mathrm{train}}$ and $\mathcal{D}_{\mathrm{val}}$ into a single
development set and hold out $\mathcal{D}_{\mathrm{test}}$.

To ask whether the provenance of the \emph{training} labels changes fairness, we
train three models that differ only in the labels they see
(\autoref{fig:regimes}): the mixed model $M_{\mathrm{mix}}$ on the full training
split, and $M_{\mathrm{gold}}$ / $M_{\mathrm{silver}}$ on its gold-only /
silver-only exams. $M_{\mathrm{mix}}$ is the deployment-realistic model --
combining scarce expert labels with abundant automated ones is the standard
response to annotation cost~\cite{tajbakhsh2020embracing}. Each regime is
sex-balanced independently, so the counts do not sum ($288 + 450 \neq 798$):
re-balancing within the gold pool alone discards its $30$-exam female surplus, and
the silver pool its male surplus. Two caveats carry into the comparison: the gold
set is smaller ($288$ vs.\ $450$) simply because fewer exams were expertly
annotated; and the gold and silver pools are demographically comparable
(``\nameref{sec:confounders}''), so label provenance is the only variable
separating $M_{\mathrm{gold}}$ from $M_{\mathrm{silver}}$.

\begin{figure}[t]
\centering
\begin{tikzpicture}
  \node[regnode=regimeCohort, text width=64mm] (D) at (0,2.5)
    {Analysis cohort $\mathcal{D}$\\ $1{,}142$ exams ($50/50$ M/F)\\[2pt]
     {\scriptsize each exam carries gold \emph{or} silver, never both}};

  \node[regnode=regimeGold,   text width=30mm] (DG) at (-2.7,0.7)
    {Gold set $\mathcal{D}^{\mathsf{g}}$\\[1pt] {\scriptsize expert annotation}};
  \node[regnode=regimeSilver, text width=32mm] (DS) at ( 2.7,0.7)
    {Silver set $\mathcal{D}^{\mathsf{s}}$\\[1pt] {\scriptsize machine-generated}};

  \node[regnode=regimeGold,   text width=22mm] (MG) at (-4.6,-1.8)
    {$M_{\mathrm{gold}}$\\[1pt] train $288$\\ {\scriptsize gold only}\\
     {\scriptsize ($50/50$ M/F)}};
  \node[regnode=regimeMix,    text width=38mm] (MM) at ( 0,  -1.8)
    {$M_{\mathrm{mix}}$\\ train $798$\\ {\scriptsize $318$ gold $+$ $480$ silver}\\
     {\scriptsize ($50/50$ M/F)}};
  \node[regnode=regimeSilver, text width=22mm] (MS) at ( 4.6,-1.8)
    {$M_{\mathrm{silver}}$\\[1pt] train $450$\\ {\scriptsize silver only}\\
     {\scriptsize ($50/50$ M/F)}};

  \draw[regedge=regimeCohort] (D)  -- (DG);
  \draw[regedge=regimeCohort] (D)  -- (DS);
  \draw[regedge=regimeGold]   (DG) -- (MG);
  \draw[regedge=regimeGold]   (DG) -- (MM);
  \draw[regedge=regimeSilver] (DS) -- (MS);
  \draw[regedge=regimeSilver] (DS) -- (MM);
\end{tikzpicture}
\caption{The three label regimes. The provenance flag $q_i$ partitions the
analysis cohort $\mathcal{D}$ into a disjoint gold (expert) set
$\mathcal{D}^{\mathsf{g}}$ and silver (machine-generated) set
$\mathcal{D}^{\mathsf{s}}$; the mixed model $M_{\mathrm{mix}}$ trains on both,
while $M_{\mathrm{gold}}$ and $M_{\mathrm{silver}}$ each train on one. Counts are
training exams and do not add up across regimes ($798 \neq 288 + 450$) because
each regime is sex-balanced independently; all three share the architecture,
training recipe, patient-level split, and stratification.}
\label{fig:regimes}
\end{figure}

\subsection{Models, Metrics, and Experimental Design}
\label{sec:model}

\paragraph{Model and training.}
We segment every regime with nnU-Net v2 (ResEnc-L
preset)~\cite{isensee2021nnunet,isensee2024revisited}, which derives its
architecture and training recipe from a dataset fingerprint. This serves the audit
twice over: it is a strong equal-resource baseline, so any gap reflects the data
and labels rather than a weak model, and holding the fingerprint-derived
configuration fixed guarantees the three regimes share one recipe -- the assumption
behind every model-to-model comparison here. Per regime we train the standard 2D
and 3D full-resolution configurations as five-fold cross-validations over the
pooled development set, and take the softmax-averaged ten-fold ensemble as the
predictor (thirty folds total). The one audit-driven change is demographically
stratified CV folds; post-processing keeps only a global largest-foreground
cleanup. The thirty folds cost ${\sim}1{,}105$ GPU-hours
(${\sim}53$--$74$\,kg\,CO$_2$e)~\cite{lacoste2019quantifying,dea_keyfigures2023}.
% Double-blind: do NOT add the HF model-release URL here for review; camera-ready/arXiv only.

\paragraph{Metrics and statistical inference.}
\label{sec:metrics}
For a foreground class $c \in \{1,2\}$ on exam $i$, with true/false positives and
false negatives $\mathrm{TP}, \mathrm{FP}, \mathrm{FN}$ between prediction and
reference, $\mathrm{Dice} = 2\,\mathrm{TP} / (2\,\mathrm{TP} + \mathrm{FP} +
\mathrm{FN})$. Because Dice's false-positive penalty is not scaled by structure
size -- the volume bias of ``\nameref{sec:confounders}'' -- we also report the
normalised Dice score (nDSC) of Raina et al.~\cite{raina2023ndsc}, which reweights
only the false-positive term by
$\kappa_{i,c} = (1 - r_c)\,|R_{i,c}| / \big(r_c\,(|\Omega_i| - |R_{i,c}|)\big)$,
where $r_c$ is the dataset-mean fraction of voxels in class $c$:
\begin{equation}
  \mathrm{nDSC}_{i,c}
  = \frac{2\,\mathrm{TP}}{\kappa_{i,c}\,\mathrm{FP} + 2\,\mathrm{TP} + \mathrm{FN}},
  \label{eq:ndsc}
\end{equation}
so each exam is scored as though its class occupied the common load $r_c$
($\kappa = 1$ recovers Dice). We add HD95 (95th-percentile Hausdorff distance, mm;
lower is better)~\cite{taha2015metrics}, all three reported per class and
macro-averaged. For group disparities we use the binarized, rate-based metrics
canonical in this
literature~\cite{parikh2025labelbias,bird2020fairlearn,eeoc1978uniform}: an exam
\emph{succeeds} when its score clears $\tau$ ($0.8$ for Dice/nDSC, $5$\,mm for
HD95), giving per-group success rates $\rho_a$, the demographic parity difference
$\mathrm{DPD} = \max_a \rho_a - \min_a \rho_a$ and disparate impact ratio
$\mathrm{DIR} = \min_a \rho_a / \max_a \rho_a$ ($<0.8$ flags adverse impact), read
off a sweep of $\tau$. For significance we test the continuous scores --
Mann--Whitney $U$~\cite{mann1947test} (sex, race) with $r_{rb}$ and
Kruskal--Wallis~\cite{kruskal1952ranks} (age) with $\varepsilon^2$, FDR-corrected
across the attribute~$\times$~metric family~\cite{benjamini1995fdr} -- backed by
BCa bootstrap CIs on each DIR, an OLS of macro Dice on sex, race, and age jointly
(adjusting the age--race entanglement of ``\nameref{sec:confounders}''), and
permutation checks. These continuous tests are our \emph{primary} instrument: in
this high-overlap regime (macro Dice ${\approx}0.89$) success rates sit near the
ceiling and DIR is near-degenerate, so the rank tests carry more power.

\paragraph{Experimental design (E1--E4).}
\label{sec:design}
\autoref{tab:experiments} lays out four experiments. E1 audits the deployed
$M_{\mathrm{mix}}$ on the full $228$-case test set, each exam scored against its own
label. E2--E4 share a common surface -- the $76$ gold-labelled test cases scored
against expert labels -- so any difference among them is attributable to the ruler
or the training labels, not the data. E2 needs two references for the same images,
but CSpineSeg gives gold or silver, never both; we generate the second by running
$M_{\mathrm{gold}}$ (trained only on our gold split, hence blind to these $76$) and
taking its predictions as the silver ruler -- mirroring how Zhou et al.\ produced
the dataset's silver labels~\cite{zhou2025cspineseg}, whose model we cannot reuse
(private weights, and a split that may overlap our test cases). Scoring
$M_{\mathrm{mix}}$'s \emph{same} predictions against the gold ruler and the
generated-silver ruler isolates the pure ruler effect. E3
($M_{\mathrm{gold}}$ vs.\ $M_{\mathrm{silver}}$) and E4 ($M_{\mathrm{mix}}$ vs.\
$M_{\mathrm{gold}}$) instead vary the training labels, testing whether silver
supervision amplifies bias or whether mixing it in harms fairness. Every experiment
applies the full metric and inference battery to every attribute in $\mathcal{A}$.

\begin{table}[t]
\centering
\caption{The four experiments. E1 is the global audit on the full test set; E2--E4
share the same $76$-case gold test set and expert reference labels, so differences
across them are attributable to the ruler or the training labels alone.}
\label{tab:experiments}
\small
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l l l l l}
\hline\noalign{\smallskip}
 & Question & Model(s) & Test set & Ruler \\
\noalign{\smallskip}\hline\noalign{\smallskip}
E1 & Is the deployed model fair? & $M_{\mathrm{mix}}$ & 228 (full) & mixed \\
E2 & Does the ruler distort the verdict? & $M_{\mathrm{mix}}$ & 76 (gold) & gold vs.\ gen.\ silver \\
E3 & Does silver training amplify bias? & $M_{\mathrm{gold}}$ vs.\ $M_{\mathrm{silver}}$ & 76 (gold) & gold \\
E4 & Does mixing silver hurt fairness? & $M_{\mathrm{mix}}$ vs.\ $M_{\mathrm{gold}}$ & 76 (gold) & gold \\
\noalign{\smallskip}\hline
\end{tabular*}
\end{table}

\paragraph{Encoder probing.}
\label{sec:probing_method}
The audit measures outputs, not representations, and encoding a protected attribute
need not produce a disparity~\cite{petersen2023feature}. We probe two frozen
encoders -- $M_{\mathrm{mix}}$'s 3D bottleneck ($320$-d) and the task-agnostic
MRI-CORE~\cite{dong2025mricore} (a ViT-B DINOv2-pretrained on ${\sim}110{,}000$ Duke
MRI volumes, cervical spine excluded, $768$-d) -- with PCA($50$) and
$\ell_2$-regularised logistic regression in five-fold cross-validation (AUROC or
balanced accuracy), each against a random-init null of identical architecture so
the trained-minus-null gap isolates learned signal from pixel-level confounds.
