\section{Discussion}
\label{sec:discussion}

The experiments above establish three factual results: (i)~the deployed model
$M_{\mathrm{mix}}$ is fair across sex, race, and age by every practical measure;
(ii)~replacing the expert reference with a machine-generated one flips the
statistical significance verdict for age; and (iii)~training on silver labels does
not amplify demographic bias. We now interpret these findings, situate them
relative to prior work, and draw out their methodological implications.

\subsection{Two Modes of Ruler Bias}
\label{sec:ruler_modes}

The verdict flip of ``\nameref{sec:biased_ruler}'' does not arise because the
silver ruler reveals a larger gap -- it does the opposite. The silver ruler is
$M_{\mathrm{gold}}$'s predictions, so scoring against it measures inter-model
agreement rather than accuracy. That agreement sits at ${\sim}0.97$ Dice -- far
above the ${\sim}0.89$ that the same predictions of $M_{\mathrm{mix}}$ achieve
against the expert labels -- and the standard
deviation across the $76$ cases drops from $0.058$ (gold) to $0.014$ (silver), a
${\sim}4\times$ reduction. We attribute this high agreement to the inherent
simplicity of cervical-spine segmentation -- vertebral bodies and discs are large,
high-contrast structures with well-defined boundaries, as evidenced by the high
Dice scores routinely achieved with standard architectures on this
anatomy~\cite{zhou2025cspineseg} -- combined with the shared architecture and
training recipe, which lead both models to converge on similar solutions
regardless of the specific training
cases.\footnote{$M_{\mathrm{mix}}$ and $M_{\mathrm{gold}}$ do share $288$
training cases, but this is not the primary driver: even $M_{\mathrm{gold}}$ and
$M_{\mathrm{silver}}$, trained on fully disjoint data (zero case overlap), still
agree at $0.97$ macro Dice on the same $76$ images. The high inter-model agreement
is a property of the task, not of shared training data.}

The Kruskal--Wallis test's power scales with signal-to-noise, so the same
clinically negligible age gradient that is buried in gold's noise clears FDR
correction once silver's collapsed variance sharpens the ranks. The ruler did not
change the patient; it changed the test's power.

Parikh et al.~\cite{parikh2025labelbias} demonstrate a complementary mode in
breast-MRI segmentation: their silver ruler is an independently trained model that
is itself biased against younger patients, so it compounds the deficit where the
patient is hard -- the performance gap between demographic groups inflates
(DIR drops from $0.871$ to $0.815$). Our silver ruler does something different: it
shrinks the between-group gap -- the continuous Dice spread falls from
$2.7$ to $0.6$ points and the rate-based DIR saturates at $1.0$ -- yet makes the
residual gradient statistically significant by collapsing within-group variance.
In their case, the ruler exaggerates the signal; in ours, it removes the noise
that would otherwise mask it. Both lead to the same practical problem -- the
fairness verdict depends on which reference label you choose -- but through
opposite routes: false magnitude versus false confidence.

\subsection{The Age Effect Is Intrinsic, Not Unfairness}
\label{sec:age_intrinsic}

Although the significance on the silver ruler is a methodological artifact, the
underlying age trend is not. It appears on both rulers in the same direction
(60+ worst), is consistent with the degenerative anatomy documented in
``\nameref{sec:confounders}'' (vertebral merging in older spines), and is
independently confirmed by the encoder probe
(``\nameref{sec:probing_results}''), which shows age encoded above the
random-init null while race sits at chance. The model has learned age-related
anatomical features -- an effect visible through both rulers and the learned
representation is unlikely to be a label artifact; it reflects intrinsic
difficulty in older spines.

Crucially, a statistically significant gap is not the same as an unfair model.
The magnitudes here are clinically negligible: the worst-vs-best age group
median spans $2.7$ Dice points on gold ($0.925 \to 0.898$) and $0.6$ points on silver, with
median performance in every age group exceeding $0.89$ Dice. All gold-ruler DIRs
exceed $0.93$ -- far above the four-fifths threshold -- and no subgroup is failed
by any practical measure. The contribution of the biased-ruler experiment is
therefore methodological: it demonstrates that a machine-generated reference label
can manufacture a finding out of a non-issue, and that significance alone should
not be read as evidence of harm.

\subsection{Why Silver Training Labels Did Not Amplify Bias}
\label{sec:no_amplification}

The null result of ``\nameref{sec:amplification}'' is expected: bias amplification
requires both an underlying disparity for noisy labels to exaggerate and a silver
source that is differentially biased across groups. Neither precondition holds
here. The global audit already found no demographic gap to amplify ($0/63$, all
DIRs $> 0.96$), and the silver labels are high-quality -- Zhou et
al.~\cite{zhou2025cspineseg} selected which exams to expert-annotate
pseudo-randomly (by medical record number), leaving the gold and silver pools
demographically comparable, so the machine-generated labels carry no systematic
group-level bias for training to propagate.

This contrasts with Parikh et al.~\cite{parikh2025labelbias}, who observed $66\%$
gap widening (DIR dropping below $0.80$) when training on their silver labels.
The difference lies in the preconditions: their silver annotator was an
independently trained model that was itself biased against younger patients,
acting on a task with a substantial pre-existing age disparity. In our setting,
the silver labels are produced by a model trained on a representative subset of
the same institution's gold annotations, segmenting large high-contrast structures
whose boundaries may leave little room for demographically differential error.
Whether it is task geometry specifically -- rather than dataset balance or
institutional homogeneity -- that provides this protection remains an open
question; isolating that factor would require a contrastive experiment on a harder
anatomy, which we leave to future work. The result suggests that mixing
silver labels into training is unlikely to harm fairness when the underlying
labels are well-calibrated and no pre-existing disparity exists for noisy
supervision to exaggerate.

This null result is necessary but not sufficient for concluding that silver labels
are universally safe. It holds here because the task is geometrically simple and
the silver source is itself unbiased; the conclusion should not be
over-generalised to settings with more complex anatomy, smaller structures, or
silver annotators of unknown provenance.

\subsection{Encoding Does Not Imply Disparity}
\label{sec:encoding_disparity}

The probe results reveal a clear dissociation: sex is by far the most strongly
encoded attribute ($0.957$ AUROC), yet the global audit found zero sex disparity
($0/63$ significant tests, DIR~$= 0.991$). The segmentation task has learned to
exploit sex-linked vertebral morphology -- larger bodies and discs in men,
different canal geometry -- without producing differential outcomes. Conversely,
age is only moderately encoded ($0.642$) yet is the one axis that shows a
(clinically negligible) performance trend, corroborating the biased-ruler finding
that the age effect is intrinsic anatomical difficulty rather than a label
artefact.

The probe thus confirms that encoding is neither necessary nor sufficient for
disparity: a model can strongly represent a demographic feature and still be fair,
because the decoder maps those different representations to equally accurate
outputs. This has direct implications for debiasing strategies that attempt to
remove demographic signal from learned representations~\cite{petersen2023feature}
-- in our setting such interventions would be unnecessary at best, and could
degrade segmentation quality by discarding morphologically informative features.

Race deserves separate attention: both the nnU-Net encoder and the MRI-CORE
foundation model encode race at chance level (AUROC ${\sim}0.50$), consistent
with the absence of racial disparity in segmentation outcomes. This contrasts
with Gichoya et al.~\cite{gichoya2022reading}, who found race highly predictable
(AUC $>0.90$) from radiographs across multiple body regions including the
cervical spine. Several differences may
explain the discrepancy: (i)~modality -- MRI and radiographs capture different
physical signals; (ii)~field of view -- our MRI volumes are tightly cropped to
the spinal column, whereas their radiographs image broader anatomy; and (iii)~training
objective -- Gichoya et al.\ trained end-to-end for race classification, whereas
our probes are linear heads on frozen encoders -- and even the task-agnostic
MRI-CORE encoder, which was not optimised for segmentation, fails to encode race.
The absence of race encoding in our MRI features
suggests that demographic leakage is not inevitable in spine imaging but depends
on the modality and the training objective.

\subsection{Limitations}
\label{sec:limitations}

We note the following limitations.

\begin{enumerate}
\item \textbf{High-performance ceiling.}
All groups exceed $0.89$ Dice and all DIRs surpass $0.93$, so there is little
room for disparity to appear. The fair verdict may reflect task simplicity rather
than a general property of the model.

\item \textbf{Single institution, anatomy, and sequence.}
All data come from Duke University, cover cervical spine only, and use one MRI
sequence (sagittal T2). There is no external validation set.

\item \textbf{Correlated silver ruler.}
CSpineSeg provides a gold or silver label per image, never both. Our generated
silver ruler is a near-clone of the model under test, so we can only show the
false-confidence effect, not the false-magnitude effect.

\item \textbf{Training-set size confound.}
The three models train on different amounts of data ($M_{\mathrm{gold}}$: $288$,
$M_{\mathrm{silver}}$: $450$, $M_{\mathrm{mix}}$: $798$). When
$M_{\mathrm{gold}}$ shows the lowest DIRs, this could reflect fewer training
cases rather than gold-label provenance. The amplification comparison cannot fully
separate label type from dataset size.

\item \textbf{Small gold test set.}
Only $76$ gold-labelled test cases, with ${\sim}20$ Black patients, limiting
power for race comparisons.

\item \textbf{Single architecture.}
All models use nnU-Net (ResEncUNetL). Other architectures may behave differently.
\end{enumerate}

\subsection{Future Work}
\label{sec:future_work}

The most direct extension is to replace our correlated silver ruler with an
independent one -- either a model trained on external data (e.g.\
TotalSpineSeg~\cite{deLeener2024totalspineseg}) or a deliberately biased
generator (e.g.\ trained only on young
patients). The former would test whether the false-confidence effect persists
when the ruler is not a twin; the latter would additionally demonstrate the
false-magnitude effect in spine and reveal how the two modes relate as ruler
correlation decreases. Second, the $4\times$ variance collapse and high
inter-model agreement that characterise our correlated ruler are detectable
signatures; formalising them into a diagnostic test would give practitioners a
way to flag when a silver ruler is too similar to the model under evaluation
before drawing fairness conclusions. Looking further ahead, Zhou et
al.~\cite{zhou2025cspineseg} note that CSpineSeg's label set may be extended to
additional structures (e.g.\ spinal cord, vertebral level classification);
repeating the fairness audit on smaller, harder-to-segment targets within the
same images and demographics would clarify whether our fair verdict reflects the
model or simply the high-performance ceiling of vertebral body and disc
segmentation.

\subsection{Recommendations for Practitioners}
\label{sec:recommendations}

We distil four guidelines from our experiments.

\begin{enumerate}
\item \textbf{State where your reference labels come from.}
  Swapping expert labels for model predictions flipped our result from $0/63$ to
  $11/63$ significant tests -- same model, same patients, same predictions.
  Without knowing the ruler's provenance, a reader cannot tell whether a reported
  finding reflects a real disparity or a measurement artifact.

\item \textbf{Check how similar the ruler is to the model.}
  In our experiment, model--ruler agreement (${\sim}0.97$ Dice) far exceeded
  model--expert agreement (${\sim}0.89$), and within-group variance dropped
  $4\times$. That collapse made a $0.6$-point age gradient clear statistical
  significance. A simple check: if scores against the ruler are much higher and
  much tighter than scores against expert labels, the ruler may be too close to
  the model to serve as an independent benchmark.

\item \textbf{Report how large the gap is, not just whether it is significant.}
  The age effect our silver ruler flags spans $0.6$ Dice points where every group
  exceeds $0.97$. Significance tells you the gap is non-zero; it does not tell
  you it matters. Always report the magnitude alongside the $p$-value.

\item \textbf{Do not remove demographic signal without an output disparity.}
  Sex is strongly encoded in our model ($0.957$ AUROC) yet produces no
  performance gap (DIR~$= 0.991$, $0/63$ significant). The model uses sex-linked
  morphology to segment well, not to discriminate. Suppressing demographic
  features~\cite{petersen2023feature} without first confirming an output disparity
  addresses a problem that may not exist.
\end{enumerate}

