\section{Methodology}
\label{sec:methodology}

% =============================================================================
% SECTION SKELETON ONLY -- no prose yet. Each subsection below lists what it
% must contain. Agreed structure: 4.1 splits/sets -> 4.2 label regimes ->
% 4.3 model/training -> 4.4 metrics -> 4.5 experimental design -> 4.6 probing.
% Register: match the formal style of the Dataset section (notation table,
% explicit definitions). Interpretation (saturation, false-confidence, "which
% ruler won") is DEFERRED to Results/Discussion -- this section is design only.
% =============================================================================

Our experimental methodology builds on the label-bias and biased-ruler framework
that Parikh et al.~\cite{parikh2025labelbias,parikh2025whofails} developed for
breast DCE-MRI, carried over here to cervical-spine segmentation. We adopt
several of their design choices -- balancing the cohort by downsampling,
comparing expert (gold) against machine-generated (silver) reference labels, and
quantifying performance gaps with binarized rate-based fairness metrics -- and we
note throughout where the structure of CSpineSeg forces us to adapt them, chiefly
that each exam carries a gold or a silver label but never both.

\subsection{Data Splits and Label Sets}
\label{sec:splits}

The Dataset section left two pieces of notation undefined, marked $(\ast)$ in
\autoref{tab:notation}: the label sets and the train/validation/test split. We
close them here, and in doing so account for the drop from the working set
$\mathcal{D}_0$ ($N_0 = 1{,}254$ exams, whose composition is summarised in
\autoref{tab:cohort}) to the analysis cohort $\mathcal{D}$ ($N = 1{,}142$), the
two cohorts introduced in \autoref{tab:notation}.

Every exam carries exactly one reference segmentation, so the provenance flag
$q_i$ partitions the cohort into two disjoint sets,
\begin{equation}
  \mathcal{D}^{\mathsf{g}} = \{\, i : q_i = \mathsf{g} \,\},
  \qquad
  \mathcal{D}^{\mathsf{s}} = \{\, i : q_i = \mathsf{s} \,\},
  \qquad
  \mathcal{D}^{\mathsf{g}} \cap \mathcal{D}^{\mathsf{s}} = \varnothing,
\end{equation}
the expert (gold) set and the machine-generated (silver) set, with
$\mathcal{D} = \mathcal{D}^{\mathsf{g}} \cup \mathcal{D}^{\mathsf{s}}$. That an
image carries gold or silver but never both is a structural property of
CSpineSeg, and it shapes the ``\nameref{sec:design}''; we
return to it in ``\nameref{sec:regimes}''.

We partition the cohort at the patient level into training, validation, and test
sets $\mathcal{D}_{\mathrm{train}}, \mathcal{D}_{\mathrm{val}},
\mathcal{D}_{\mathrm{test}}$ in a $70/10/20$ ratio. Splitting by patient rather
than by exam keeps all of a patient's exams on the same side of every boundary,
so the handful of multi-exam patients cannot leak information from training into
evaluation. The split is stratified by the protected attributes $\mathcal{A}$
so that each subgroup's share stays roughly constant across
the three sets.\footnote{White and Black patients, who form the race comparison,
are stratified on the full race $\times$ age $\times$ sex key; the small
remaining race groups (\autoref{tab:cohort}) are too sparse to stratify and are
assigned to splits proportionally at random.}

On top of this we balance the sexes. The working set leans female
(\autoref{tab:cohort}), and an unequal base rate would confound the sex audit by
conflating a true performance gap with mere differences in representation, so
within each split we randomly downsample female exams -- the over-represented sex
throughout this cohort -- until the two are equal, removing $112$ in total
($N_0 = 1{,}254 \to N = 1{,}142$) and leaving every split exactly $50/50$
male/female. This removes representation as an explanation for any residual gap,
rather than attempting to close it~\cite{parikh2025labelbias}. The downsampling is seeded
and reproducible, and because it operates at the exam level while the
patient-level split already fixes which side each patient falls on, it introduces
no leakage even for the multi-exam patients. We adopt this balance-by-downsampling
protocol from Parikh et al.~\cite{parikh2025labelbias,parikh2025whofails}, who
balance their breast-MRI cohort by age to probe representational imbalance; we
apply the same idea to sex and to this dataset's multi-exam structure.
Stratification held in every split: race and age proportions match the cohort
marginals of \autoref{tab:cohort} to within a couple of percentage points, and
sex is exactly $50/50$ by construction. The
per-regime breakdown of these splits into gold and silver training cases is given
in ``\nameref{sec:regimes}''.

We keep the conventional $70/10/20$ split rather than a bare train/test partition, which
keeps the cohort partition portable across modelling choices. nnU-Net is
cross-validation-native -- it selects its configuration and post-processing through an
internal five-fold cross-validation rather than a held-out tuning
set~\cite{isensee2021nnunet} -- so for our segmentation models $\mathcal{D}_{\mathrm{train}}$
and $\mathcal{D}_{\mathrm{val}}$ are pooled into the single development set that the
cross-validation runs over, leaving none of the scarce expert annotations idle (most
valuable in the gold-only regime, ``\nameref{sec:regimes}''). The validation tier still
earns its place: a model that is not cross-validation-native -- a transformer baseline, or
any architecture used to extend this audit -- needs a held-out set for early stopping and
hyperparameter selection, so defining $\mathcal{D}_{\mathrm{val}}$ now lets such work tune
and report on the identical partition while $\mathcal{D}_{\mathrm{test}}$ remains the single
fully held-out test set for every method.

\subsection{Three Label Regimes}
\label{sec:regimes}

We want to know whether the provenance of the training labels -- expert
or machine-generated -- changes the fairness of the resulting model. To isolate
that effect we train three models that differ only in the labels they are given,
holding everything else fixed: the same architecture and training recipe
(``\nameref{sec:model}''), the same patient-level split of
``\nameref{sec:splits}'', and the same sex balance and stratification. Any
difference in their fairness behaviour is then attributable to the training
labels alone, and not to model capacity or cohort composition -- the assumption
that justifies every model-to-model comparison in this paper.

The three regimes follow directly from the disjoint gold/silver partition
$\mathcal{D} = \mathcal{D}^{\mathsf{g}} \cup \mathcal{D}^{\mathsf{s}}$ introduced
in ``\nameref{sec:splits}''. The mixed model $M_{\mathrm{mix}}$ trains on the
full training split $\mathcal{D}_{\mathrm{train}}$, taking each exam with
whichever label it carries; combining a small pool of expert labels with a
larger pool of automated ones is a standard response to the cost of expert
annotation in medical image segmentation~\cite{tajbakhsh2020embracing}, so
$M_{\mathrm{mix}}$ is the deployment-realistic model that a typical pipeline
would train. The
gold model $M_{\mathrm{gold}}$ and the silver model $M_{\mathrm{silver}}$
restrict training to only the expert or only the machine-generated exams, drawn
from $\mathcal{D}_{\mathrm{train}} \cap \mathcal{D}^{\mathsf{g}}$ and
$\mathcal{D}_{\mathrm{train}} \cap \mathcal{D}^{\mathsf{s}}$ respectively.
\autoref{fig:regimes} shows how the three regimes are drawn from the gold/silver
partition, and \autoref{tab:regimes} gives their full composition.

That the gold-only and silver-only regimes can be built at all rests on the
structural property of CSpineSeg flagged above: each exam carries a gold or a
silver label, never both ($\mathcal{D}^{\mathsf{g}} \cap \mathcal{D}^{\mathsf{s}}
= \varnothing$). This single-label structure is the chief way our setting departs
from the breast-MRI biased-ruler work we build on, and it shapes the experimental
design (``\nameref{sec:design}'').

\begin{figure}[t]
\centering
\begin{tikzpicture}
  \node[regnode=regimeCohort, text width=64mm] (D) at (0,2.5)
    {Analysis cohort $\mathcal{D}$\\ $1{,}142$ exams ($50/50$ M/F)\\[2pt]
     {\scriptsize each exam carries gold \emph{or} silver, never both}};

  \node[regnode=regimeGold,   text width=30mm] (DG) at (-2.7,0.7)
    {Gold set $\mathcal{D}^{\mathsf{g}}$\\[1pt] {\scriptsize expert annotation}};
  \node[regnode=regimeSilver, text width=32mm] (DS) at ( 2.7,0.7)
    {Silver set $\mathcal{D}^{\mathsf{s}}$\\[1pt] {\scriptsize machine-generated}};

  \node[regnode=regimeGold,   text width=22mm] (MG) at (-4.6,-1.8)
    {$M_{\mathrm{gold}}$\\[1pt] train $288$\\ {\scriptsize gold only}\\
     {\scriptsize ($50/50$ M/F)}};
  \node[regnode=regimeMix,    text width=38mm] (MM) at ( 0,  -1.8)
    {$M_{\mathrm{mix}}$\\ train $798$\\ {\scriptsize $318$ gold $+$ $480$ silver}\\
     {\scriptsize ($50/50$ M/F)}};
  \node[regnode=regimeSilver, text width=22mm] (MS) at ( 4.6,-1.8)
    {$M_{\mathrm{silver}}$\\[1pt] train $450$\\ {\scriptsize silver only}\\
     {\scriptsize ($50/50$ M/F)}};

  \draw[regedge=regimeCohort] (D)  -- (DG);
  \draw[regedge=regimeCohort] (D)  -- (DS);
  \draw[regedge=regimeGold]   (DG) -- (MG);
  \draw[regedge=regimeGold]   (DG) -- (MM);
  \draw[regedge=regimeSilver] (DS) -- (MS);
  \draw[regedge=regimeSilver] (DS) -- (MM);
\end{tikzpicture}
\caption{The three label regimes. The provenance flag $q_i$ partitions the
analysis cohort $\mathcal{D}$ into a disjoint gold (expert) set
$\mathcal{D}^{\mathsf{g}}$ and silver (machine-generated) set
$\mathcal{D}^{\mathsf{s}}$; the mixed model $M_{\mathrm{mix}}$ trains on both,
while $M_{\mathrm{gold}}$ and $M_{\mathrm{silver}}$ each train on one. Counts are
training exams and do not add up across regimes ($798 \neq 288 + 450$) because
each regime is sex-balanced independently; the full per-split composition is in
\autoref{tab:regimes}.}
\label{fig:regimes}
\end{figure}

\begin{table}[t]
\centering
\caption{The three label regimes. All three share the architecture, training
recipe, patient-level split, stratification, and exact $50/50$ sex balance, and
differ only in the provenance of their training labels. Counts are exams; every
split in every regime is exactly $50/50$ male/female. Validation and test counts
are regime-specific and are not comparable across rows.}
\label{tab:regimes}
\small
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l l r r r}
\hline\noalign{\smallskip}
Model & Training labels & Train & Val & Test \\
\noalign{\smallskip}\hline\noalign{\smallskip}
$M_{\mathrm{mix}}$    & gold $+$ silver & 798 (318 gold $+$ 480 silver) & 116 & 228 \\
$M_{\mathrm{gold}}$   & gold only       & 288                            &  44 &  76 \\
$M_{\mathrm{silver}}$ & silver only     & 450                            &  66 & 138 \\
\noalign{\smallskip}\hline
\end{tabular*}
\end{table}

Because the gold and silver sets are disjoint, each regime is sex-balanced on
its own, and the counts do not add up across regimes. The principle is that a
set can be exactly $50/50$ male/female while neither of its two parts is. The
mixed training split is balanced across all $798$ of its exams ($399$ female and
$399$ male), but that balance is a cancellation: its $318$ gold exams lean female
($174$ versus $144$) while its $480$ silver exams lean male ($255$ versus $225$),
so the $30$-exam female surplus on the gold side exactly offsets the $30$-exam
male surplus on the silver side. Re-balancing sex within the gold pool alone
therefore drops those $30$ surplus female exams, which is why $M_{\mathrm{gold}}$
trains on $288$ gold exams ($144$ of each sex) rather than the $318$ that appear
in the mixed split -- a balanced subset of the very same cases, not a different
draw. The silver regime mirrors it, dropping the $30$ surplus male exams to reach
$450$ ($225$ of each sex). The regimes consequently do not sum --
$288 + 450 \neq 798$ -- because each discards its own surplus sex independently.
Validation and test counts are likewise regime-specific (\autoref{tab:regimes})
and should not be compared across rows.

One asymmetry remains by construction: the gold training set ($288$ exams) is
smaller than the silver one ($450$), simply because fewer CSpineSeg exams were
expertly annotated in the first place. This reflects the dataset's structure
rather than a design choice, and we flag it as a limitation of the
gold-versus-silver comparison.

A second concern is subtler. Because $M_{\mathrm{gold}}$ and $M_{\mathrm{silver}}$
train on disjoint exams, a demographic imbalance between the gold and silver sets
would reintroduce cohort composition as an explanation for any fairness difference
between them. It does not arise here: Zhou et al.~chose which exams to
expert-annotate pseudo-randomly -- by medical-record number rather than by any
demographic criterion~\cite{zhou2025cspineseg} -- so the gold and silver sets are
demographically comparable, and label provenance is the only variable that
separates them.

\subsection{Model and Training}
\label{sec:model}

We segment every regime with nnU-Net~\cite{isensee2021nnunet}, the self-configuring
framework that has become the standard baseline for biomedical segmentation. Rather
than fix an architecture by hand, nnU-Net derives the whole pipeline -- patch and
batch size, network depth, resampling, and intensity normalisation -- from a
fingerprint of the dataset (image sizes, voxel spacings, and intensity
statistics) through a set of rule-based heuristics. Two properties make it the right
backbone for a fairness audit. First, it is a genuinely strong baseline: under
controlled, equal-resource validation a well-scaled nnU-Net still matches or exceeds
the transformer and state-space architectures that claim to surpass
it~\cite{isensee2024revisited}, so any performance gap we report is a property of the
data and its labels, not an artefact of a weak model. Second, because the
configuration is derived from the fingerprint rather than chosen per run,
holding the framework fixed across the three regimes guarantees that
$M_{\mathrm{mix}}$, $M_{\mathrm{gold}}$, and $M_{\mathrm{silver}}$ share an identical
architecture and training recipe -- the assumption that justifies every model-to-model
comparison in ``\nameref{sec:regimes}''.

We use nnU-Net v2 with the residual-encoder ``L'' preset (ResEnc-L), the framework's
current recommended default, which devotes a larger GPU-memory budget to a
residual-encoder U-Net~\cite{isensee2024revisited}. Because CSpineSeg is MRI, the planner
selects per-case $z$-score intensity normalisation. Following nnU-Net's standard protocol
we train two configurations for every regime -- a 2D model and a full-resolution 3D model
(patches $512\times512$ and $16\times512\times512$, the latter's shallow depth matching
the thin, few-slice sagittal stacks of CSpineSeg) -- each as a five-fold
cross-validation over the pooled development set
($\mathcal{D}_{\mathrm{train}}\cup\mathcal{D}_{\mathrm{val}}$) defined in
``\nameref{sec:splits}''. nnU-Net's own
configuration search then selects, for each regime, the softmax-averaged ensemble of
both configurations across all five folds -- ten trained models -- as the final
predictor. The same recipe runs for all three regimes, changing only the training
labels.

One choice departs from nnU-Net's defaults, to serve the audit: we replace the
framework's random cross-validation folds with demographically stratified ones, so that
each fold preserves the joint race~$\times$~age~$\times$~sex composition of the cohort;
together with the patient-level, sex-balanced split of ``\nameref{sec:splits}'' this
keeps subgroup representation constant through training and validation. Post-processing,
by contrast, we leave to nnU-Net's automatic search, which for this multi-object anatomy
declines the per-class ``keep largest component'' cleanup that its single-object default
would otherwise apply: collapsing each class to its largest blob would discard all but one
of the ${\sim}7$ vertebrae and the multiple discs (vertebral-body Dice $0.96 \to 0.49$ on
our cross-validation folds). Only a global largest-foreground cleanup, which leaves the
connected spine column intact, is retained, and it shifts the score negligibly. Training
otherwise uses the stock nnU-Net trainer; our only addition is experiment logging, leaving
the loss, optimiser, and augmentation untouched. The splits and the sex-balancing
downsample are seeded and reproducible.
% TODO: models not yet released -- add public-model link here once the Hugging Face
% repo is live (see docs/nnunet/07_huggingface_release.md). E.g.: ", and the trained
% models are released publicly at \url{...}." Do NOT claim release until it is public.
Training all thirty
model-folds consumed roughly $1{,}105$ GPU-hours
on A100 and L40S hardware; under the standard ML~CO$_2$
accounting~\cite{lacoste2019quantifying,strubell2019energy,patterson2021carbon} and the
2023 Danish grid intensity~\cite{dea_keyfigures2023} this corresponds to an estimated
$53$--$74$\,kg~CO$_2$e.

\subsection{Evaluation Metrics and Statistical Analysis}
\label{sec:metrics}

We evaluate every model along two axes: how well it segments, and how evenly that
quality is spread across the protected groups $\mathcal{A}$. The first calls for
segmentation-quality metrics, the second for group-level fairness metrics, and
reading either as evidence calls for a statistical-inference layer on top. We
define all three here; which experiment applies them to which model and labels is
deferred to ``\nameref{sec:design}''.

\paragraph{Segmentation-quality metrics.}
Fix an exam $i$ and a foreground class $c \in \{1, 2\}$ (vertebral body, disc).
Writing $\hat{Y}_i$ for the model's prediction and $Y_i$ for the reference
(\autoref{tab:notation}), the predicted and reference voxel sets for class $c$ are
$P_{i,c} = \{v \in \Omega_i : \hat{Y}_i(v) = c\}$ and
$R_{i,c} = \{v \in \Omega_i : Y_i(v) = c\}$, from which we take the usual counts
$\mathrm{TP} = |P_{i,c} \cap R_{i,c}|$, $\mathrm{FP} = |P_{i,c} \setminus R_{i,c}|$,
and $\mathrm{FN} = |R_{i,c} \setminus P_{i,c}|$. The Dice
score~\cite{dice1945measures} is the overlap
\begin{equation}
  \mathrm{Dice}_{i,c}
  = \frac{2\,\mathrm{TP}}{2\,\mathrm{TP} + \mathrm{FP} + \mathrm{FN}}.
  \label{eq:dice}
\end{equation}

Dice has the volume bias the Dataset section warned of (``\nameref{sec:confounders}''):
because the false-positive penalty is not scaled by structure size, a fixed
boundary error costs a smaller fraction of a large structure's score than of a
small one's, so the larger anatomy of men and of older patients can score higher
at equal true quality. The normalised Dice score (nDSC) of
Raina et al.~\cite{raina2023ndsc} -- the volume-aware score promised in
``\nameref{sec:confounders}'' -- removes this by reweighting only the false-positive
term. Let $r_c$ be a dataset-level reference load, the mean fraction of
voxels that class $c$ occupies,
\begin{equation}
  r_c = \frac{1}{N}\sum_{i} \frac{|R_{i,c}|}{|\Omega_i|},
  \label{eq:load}
\end{equation}
and define the per-exam weight
$\kappa_{i,c} = \dfrac{(1 - r_c)\,|R_{i,c}|}{r_c\,(|\Omega_i| - |R_{i,c}|)}$. Then
\begin{equation}
  \mathrm{nDSC}_{i,c}
  = \frac{2\,\mathrm{TP}}{\kappa_{i,c}\,\mathrm{FP} + 2\,\mathrm{TP} + \mathrm{FN}}.
  \label{eq:ndsc}
\end{equation}
The weight $\kappa_{i,c}$ rescales each exam's false-positive penalty as though its
class occupied the common load $r_c$ rather than its own size -- equivalently,
$\kappa_{i,c} = 1$ recovers Dice (\autoref{eq:dice}) exactly when an exam's
reference fraction equals $r_c$ -- so nDSC decouples the score from absolute
structure volume and neutralises the size confound. Alongside these two overlap
scores we report a boundary metric, the $95$th-percentile Hausdorff distance
(HD95, in mm)~\cite{taha2015metrics}: the $95$th percentile of the symmetric set
of surface distances between the predicted and reference boundaries of class $c$,
which captures localisation error that overlap scores miss while the percentile
trims the sensitivity of the raw Hausdorff distance to a few stray voxels (lower
is better). All three metrics are computed per class and reported separately for
vertebral bodies and discs, together with their macro average over the two
foreground classes.

\paragraph{Fairness metrics.}
To turn a per-exam score into a group-level disparity we adopt the binarized,
rate-based fairness metrics canonical in this literature -- Parikh
et al.~\cite{parikh2025labelbias,parikh2025whofails}, Fairlearn~\cite{bird2020fairlearn},
and the four-fifths rule~\cite{eeoc1978uniform}. Each exam is first reduced to a
binary beneficial outcome (a ``success''): for the higher-is-better overlap
scores the outcome succeeds when the score clears a threshold $\tau$, and for
HD95, where lower is better, when it falls below $\tau$. Following Parikh
et al.~\cite{parikh2025labelbias,parikh2025whofails} we take $\tau = 0.8$ as the
beneficial-outcome cutoff for Dice and nDSC, and set $\tau = 5$\,mm for HD95; all
thresholds are configurable. For an attribute $A$ with value $a$ the group success rate is the
share of its exams that succeed,
\begin{equation}
  \rho_a = \frac{1}{|\mathcal{D}_{A=a}|}
           \sum_{i \in \mathcal{D}_{A=a}} \mathbf{1}[\text{exam $i$ succeeds}],
  \label{eq:rate}
\end{equation}
and the two disparity measures compare these rates across the groups of $A$,
\begin{equation}
  \mathrm{DPD}_A = \max_a \rho_a - \min_a \rho_a,
  \qquad
  \mathrm{DIR}_A = \dfrac{\min_a \rho_a}{\max_a \rho_a}.
  \label{eq:dpddir}
\end{equation}
The demographic parity difference (DPD) is the absolute gap in success rate and
the disparate impact ratio (DIR) the relative one, with $\mathrm{DIR}_A = 1$
denoting parity. Following the four-fifths rule we flag $\mathrm{DIR}_A < 0.8$ as
adverse impact, a threshold that applies legitimately here precisely because DIR
is a ratio of selection (success) rates rather than of continuous means. The rule
does carry a portability caveat: it originates in US employment
law~\cite{eeoc1978uniform}, and Fairlearn~\cite{bird2020fairlearn} notes it has no
validity outside that setting, so we adopt the $0.8$ line as the established
convention of the fairness-audit literature and for direct comparability with
Parikh et al., not as a legal claim. Because a single threshold can be brittle, we
additionally sweep DPD and DIR over a grid of $\tau$ with every run and
read the verdict off the full sweep rather than any one cutoff.

\paragraph{Statistical inference.}
The fairness metrics summarise the size of a gap; establishing that one is real
calls for inference on the underlying scores. Our primary instrument runs on the
continuous per-exam scores, using the same rank-based battery as the
Dataset section: a Mann--Whitney $U$ test~\cite{mann1947test} for the two-group
attributes (sex, and White vs.\ Black race), with the rank-biserial correlation
$r_{rb}$~\cite{wendt1972rankbiserial} as effect size, and a Kruskal--Wallis $H$
test~\cite{kruskal1952ranks} across the three age bins, with
$\varepsilon^2$~\cite{tomczak2014effectsize}. Because we test a family of
attribute~$\times$~metric combinations per model, we control the false discovery
rate across that family with the Benjamini--Hochberg
procedure~\cite{benjamini1995fdr} and report raw and adjusted $p$-values. Three
further layers support this. We attach BCa bootstrap $95\%$ confidence
intervals~\cite{efron1987bca} to each DIR, so the disparity estimates carry
uncertainty rather than standing as point values. Distinct from the anatomy-driven volume confound that nDSC already neutralises,
one protected attribute can also confound another: ``\nameref{sec:confounders}''
flagged that White patients are slightly older, so age can leak into the race
comparison. This is the sole such entanglement here -- sex is balanced and
independent of the other attributes -- and we control for it by additionally
fitting a multivariable ordinary-least-squares regression of macro Dice on sex,
race, and age jointly, so each attribute's association is estimated adjusted for
the others rather than marginally. And we corroborate the rank-based
$p$-values with distribution-free permutation tests, which make no asymptotic
assumption about the bounded, left-skewed score distributions.

\paragraph{Which instrument is primary.}
The continuous tests, not the binarized DIR, are our primary fairness instrument,
and the reason is structural rather than empirical. Our models operate in a
high-overlap regime -- macro Dice around $0.89$ -- so the great majority of exams
clear any reasonable beneficial-outcome threshold and the per-group success rates
of \autoref{eq:rate} sit close to the ceiling. With few failures in any group the
DIR of \autoref{eq:dpddir} is near-degenerate by construction: it has little room
to move and little power to resolve a true gap. The continuous Mann--Whitney and
Kruskal--Wallis tests, which use the full distribution of scores rather than a
thresholded count, are therefore by design the more sensitive instrument in this
regime. We retain DPD and DIR for comparability with Parikh et al.\ and for the
four-fifths reading, guarded by the threshold sweep, but we treat the continuous
tests and their effect sizes as the primary evidence throughout.

\subsection{Experimental Design}
\label{sec:design}

The three regimes and the metrics above combine into four experiments, summarised
in \autoref{tab:experiments}. Each asks a distinct question about the relationship
between label provenance and fairness, and together they cover the two ways silver
labels can distort a fairness verdict: through the reference used to evaluate
a model (the ``ruler''), and through the labels used to train it.

\paragraph{Test sets.}
The first experiment (E1) uses the full $228$-case test set
$\mathcal{D}_{\mathrm{test}}$ and scores each exam against whichever reference
label it carries -- the operational setting for a realistically trained model. The
remaining three (E2--E4) all evaluate on the same $76$ gold-test cases
$\mathcal{D}_{\mathrm{test}} \cap \mathcal{D}^{\mathsf{g}}$, scored against expert
labels only. This common evaluation surface ensures that any difference between
E2--E4 is attributable to the ruler or the training labels, not to the test data.

\paragraph{The generated-silver ruler.}
E2 requires two reference segmentations for the same $76$ images, but CSpineSeg
images carry gold or silver -- never both
(``\nameref{sec:splits}''). We construct the second ruler by running
$M_{\mathrm{gold}}$ -- trained exclusively on our gold split and therefore blind to
these $76$ test cases -- on the same images and taking its predictions as
machine-generated reference labels. This mirrors how the dataset's original silver
labels were produced: Zhou et al.\ trained an nnU-Net on their expert annotations
and applied it to the unannotated cases~\cite{zhou2025cspineseg}. We cannot reuse
their model because its weights are not public and its training split may overlap
our test cases, a leakage risk; $M_{\mathrm{gold}}$, trained on our own split,
excludes them by design.

The biased-ruler comparison then scores the same predictions of
$M_{\mathrm{mix}}$ on the $76$ images twice: once against the expert labels
(Ruler~A, gold) and once against the generated-silver labels (Ruler~B). Because
$M_{\mathrm{mix}}$ and $M_{\mathrm{gold}}$ are independently trained models that
share no weights, their predictions differ and the comparison is not degenerate.
Any change in the observed fairness gap between the two rulers is therefore the
pure effect of the reference label.

\paragraph{Bias amplification and mixed-vs-gold training.}
E3 and E4 address the complementary question: does label provenance in
training -- rather than in evaluation -- change the fairness of the
resulting model? E3 compares $M_{\mathrm{gold}}$ against $M_{\mathrm{silver}}$ on
the $76$ gold-test cases, both scored against expert labels, isolating whether
silver training labels widen demographic gaps relative to gold. E4 is the
production-realistic companion: it compares $M_{\mathrm{mix}}$ against
$M_{\mathrm{gold}}$ on the same surface, asking whether mixing silver labels into
training -- the standard cost-saving practice~\cite{tajbakhsh2020embracing} --
degrades fairness relative to expert-only training.

\begin{table}[t]
\centering
\caption{The four experiments. E1 is the global audit on the full test set; E2--E4
share the same $76$-case gold test set and expert reference labels, so differences
across them are attributable to the ruler or the training labels alone.}
\label{tab:experiments}
\small
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l l l l l}
\hline\noalign{\smallskip}
 & Question & Model(s) & Test set & Ruler \\
\noalign{\smallskip}\hline\noalign{\smallskip}
E1 & Is the deployed model fair? & $M_{\mathrm{mix}}$ & 228 (full) & mixed \\
E2 & Does the ruler distort the verdict? & $M_{\mathrm{mix}}$ & 76 (gold) & gold vs.\ gen.\ silver \\
E3 & Does silver training amplify bias? & $M_{\mathrm{gold}}$ vs.\ $M_{\mathrm{silver}}$ & 76 (gold) & gold \\
E4 & Does mixing silver hurt fairness? & $M_{\mathrm{mix}}$ vs.\ $M_{\mathrm{gold}}$ & 76 (gold) & gold \\
\noalign{\smallskip}\hline
\end{tabular*}
\end{table}

All four experiments apply the full metric and inference battery of
``\nameref{sec:metrics}'' -- Dice, nDSC, HD95, DPD/DIR with threshold sweep,
Mann--Whitney/Kruskal--Wallis tests, bootstrap CIs, and the OLS confounder
regression -- to every attribute in $\mathcal{A}$. We report results per class
(vertebral body, disc) and as the macro average.

\subsection{Encoder Probing}
\label{sec:probing_method}

The fairness audit tells us whether the model's output differs across groups; it
cannot tell us whether the model's internal representation carries demographic
signal in the first place. A representation that encodes a protected attribute
could drive a downstream disparity, but it need not -- encoding and disparity are
logically independent~\cite{petersen2023feature}. We therefore complement the
audit with a linear probing diagnostic that asks, for each attribute in
$\mathcal{A}$: can it be decoded from the encoder's features?

We probe two encoders. The first is $M_{\mathrm{mix}}$'s bottleneck: the
global-average-pooled final encoder stage of the 3D ResEnc-L ($320$-d), whose
features are shaped by the segmentation objective on this cohort. The second is
MRI-CORE~\cite{dong2025mricore}, a ViT-B pretrained with DINOv2 self-supervision
on ${\sim}110{,}000$ Duke clinical MRI volumes (cervical spine excluded), which
provides a task-agnostic $768$-d reference. For each we fit PCA($50$) followed by
$\ell_2$-regularised logistic regression in five-fold stratified
cross-validation, scoring binary attributes by AUROC and multi-class ones by
balanced accuracy, with Nadeau--Bengio corrected confidence
intervals~\cite{nadeau2003inference}.

Each encoder is paired with a random-initialisation null -- the same architecture
and input pipeline with untrained weights -- which isolates learned signal from
trivial pixel-level confounds (body extent, intensity histogram). The difference
between the trained and null scores is the contribution of learning to the
demographic signal.
